---
title: "Inclass Exercise-1"
author: "Emre Çakmak"
date: "2022-10-19"
toc: true
toc-location: left
number-sections: true
format: 
  html:
    theme: 
      dark: vapor
      light: pulse
    fontsize: 0.8em
editor: visual
---

### INCLASS EXERCISE

This exercise has been prepared for understanding {dplyr} package usage for functional EDA. Main data set in this exercise will be planes data set derived from FAA.

First of all we have to install our packages.

```{r install_packages, eval=FALSE}
install.packages("tidyverse")
install.packages("nycflights13")

```

Then we are calling our libraries.

```{r call_libraries, message = FALSE}
library(tidyverse)
library(nycflights13)
```

Let's check first 10 rows of the data set. Fields and their meanings are:

-   **tailnum**: Tail number.
-   **year**: Year manufactured.
-   **type**: Type of plane.
-   **manufacturer**: Manufacturer of the aircraft.
-   **model**: Model of the aircraft.
-   **engines**: Number of engines
-   **seats**: Number of seats
-   **speed**: Average cruising speed in mph.
-   **engine**: Type of engine.

```{r define_dataset}
planes %>% 
  slice(1:10)
```

## EXERCISE 1

Now, how many aircraft does exists for each manufacturing company? Let's calculate..

```{r group_data}
planes %>% 
  group_by(manufacturer) %>% 
  summarise(aircraft_count = n()) %>% 
  arrange(desc(aircraft_count)) %>% 
  print(n=Inf)
```

It seems like there is a conflict in manufacturer names. Some of them represent the same company but in different names like Airbus and Airbus Industrie.

We need to clean and rewrite these names. Then we can apply same process again.

```{r simplify_manufacturer_names}
planes =
planes %>% 
  mutate(manufacturer = gsub("AIRBUS INDUSTRIE", "AIRBUS", manufacturer), manufacturer=gsub(" AIRCRAFT CO| CORPORATION", "", manufacturer))
```

The last version of distribution of air crafts according to their manufacturer is here.

```{r regroup_data}
planes %>% 
  group_by(manufacturer) %>% 
  summarise(aircraft_count = n()) %>% 
  arrange(desc(aircraft_count)) %>% 
  mutate(aircraft_count_distrubiton=round(aircraft_count/sum(aircraft_count),2)) %>% 
  print(n=Inf)
```

## EXERCISE 2

Let's check the difference on aircraft capacities year by year

First, get only air crafts which have more than 50 seats. Then clear the data by filtering rows which have no information in Year column.

```{r year_by_year_capacity}
planes %>% 
  filter(seats>50,!is.na(year))%>%
  group_by(year) %>% 
  summarise(seat_avg = round(mean(seats),2)) %>% 
  arrange(year) %>% 
  print(n=Inf)
```

Let's check the biggest air craft in our database with it's tailnumber.

```{r biggest_aircraft}
planes %>%
  arrange(desc(seats)) %>% 
  slice(1) %>%
  select(tailnum, manufacturer, model, year, seats)
```

Exciting..Here is some information about [the biggest airplane's history](https://www.planespotters.net/airframe/boeing-747-400-n670us-delta-air-lines/3wno03)

**THANKS FOR READING**
